---
title: "Entropy"
author: "WTHOML-tvjoseph"
date: "Monday, June 15, 2015"
output: html_document
---
Introduction

This is a program to calculate the entropy between two variables in a dataframe. The variables can be selected by the user along with the window within which entropy needs to be calculated for the given set of variables

Step:1

This program is dependent on the following library. Please run this library function to begin. In addition define the path where the csv file is located

User Action : Run the library function and define the path to the csv file

```{r}
library(entropy)

path <- "E:/elance/entropy1.csv"

```

Step :2

The below function is the function to read the csv files and to define the entropy matrix and correlation matrix.

User Action : Run the entr1 function

```{r}

entr1 <- function(path){  
  entrop <- read.csv(path) ## Reading the csv files containing the variables
  variables <- names(entrop) ## Finding the names of variables to select the variables for calculating entropy
  varrows <- nrow(entrop) ## Finding number of rows in the df
  varcol <- ncol(entrop) ## Finding the number of columns of the df
  entr.matrix1 <- matrix(0,nrow = varcol,ncol = varcol) ## Defining empty matrices 
  entr.matrix2 <- matrix(0,nrow = varcol,ncol = varcol) ## Defining empty matrices
  colnames(entr.matrix1) <- variables
  rownames(entr.matrix1) <- variables
  colnames(entr.matrix2) <- variables
  rownames(entr.matrix2) <- variables
  list(result1=entr.matrix1,result2=entr.matrix2,result3=variables,result4=varrows,result5=entrop)
}

```

Step : 3

The below steps are to assign the matrices and certain variables outside the function so that it is available within the global environment.

User Action : Run the below lines

```{r}

entr.matrix1 <- entr1(path)$result1 ## Run the Entropy matrix
entr.matrix2 <- entr1(path)$result2 ## Run the Correlation matrix
entrop <- entr1(path)$result5
varrows <- entr1(path)$result4

```

Step : 4

The below function is the main function to calculate the entropies between the selected variables and also to calculate the correlation between the variables. 

User Action : Run the entr2 function

```{r}

entr2 <- function(var1,var2,entrow1,entrow2,bin1,bin2){
  if(((entrow2-entrow1+1) > varrows)|(entrow1 > varrows)|(entrow2 > varrows)){stop("Rows selected are out of range")} ## Stopping the function if the rows selected are greater than the rows present
  var1 <- paste0(var1,"$")
  var2 <- paste0(var2,"$")
  i <- grep(var1,colnames(entrop)) ## Finding the index of the first variable
  j <- grep(var2,colnames(entrop)) ## Finding the index of the second variable
  
  X1 <- entrop[entrow1:entrow2,i] ## Selecting the first variable vector
  Y1 <- entrop[entrow1:entrow2,j] ## Selecting the second variable vector
  
    
  var.matrix <- discretize2d(X1,Y1,numBins1=bin1,numBins2=bin2) ## Making a matrix of the frequencies
  
  H1 <- entropy(rowSums(var.matrix)) ## Entropy of the first variable
  H2 <- entropy(colSums(var.matrix)) ## Entropy of the second variable
  H12 <- entropy(var.matrix) ## Joint entropy
  gencor <- H1+H2-H12 ## Finding the correlation between the variables
  
  list(result1=H12,result2=gencor,result3=i,result4=j)    
  
}


```

Step 5 

The below lines have to be executed to find the names of the variables and also to find the number of rows in the dataframe , so that the user can select the appropriate variables and sliding window.

User Action : Run the below lines and select the variables for subsequent steps

```{r}
entr1(path)$result3 ## Run this line for selecting the variables
entr1(path)$result4 ## Run this line for finding total rows in the dataframe

```

Step 6 

The below steps executes the main function to calculate the entropy and correlation between the variables. 

User Action : Enter var1, var2,window1,window2,bin1 & bin2 for executing the function.

```{r}
i <- entr2("var3","var5",6,14,5,5)$result3 ## Define the variables in the following order Variable1,variable2,window1,window2,bin1,bin2
j <- entr2("var3","var5",6,14,5,5)$result4 

entr.matrix1[i,j] <- entr2("var3","var5",6,14,5,5)$result1 ## Run this line to get the values for the entropy matrix
entr.matrix1[j,i] <- entr.matrix1[i,j] ## Run this line to get both values for the entropy matrix
entr.matrix2[i,j] <- entr2("var3","var5",6,14,5,5)$result2 ## Run this line to get the values for the correlation matrix
entr.matrix2[j,i] <- entr.matrix2[i,j] ## Run this line to get both values for the correlation matrix

entr.matrix1 ## displaying the entropy matrix
entr.matrix2 ## displaying the correlation matrix
```

