---
title: "Entropy2"
author: "WTHOML-tvjoseph"
date: "Wednesday, June 17, 2015"
output: html_document
---

Introduction
 
This is a program to calculate the entropy between two variables in a dataframe. The variables can be selected by the user along with the window within which entropy needs to be calculated for the given set of variables
 
Step:1
 
Define the path where the csv file is located
User Action : Define the path to the csv file
 
 
```{r}
path <- "E:/elance/entropy.csv"

```

 
Step :2
 
The below function is the function to read the csv files
User Action : Run the entr1 function

```{r}
entr1 <- function(path){ 
  entrop <- read.csv(path) ## Reading the csv files containing the variables
  variables <- names(entrop) ## Finding the names of variables to select the variables for calculating entropy
  varrows <- nrow(entrop) ## Finding number of rows in the df
  varcol <- ncol(entrop) ## Finding the number of columns of the df
   list(result1=entrop,result2=varrows,result3=varcol)
}

```
 
Step : 3
 
The below steps are to assign the matrices and certain variables outside the function so that it is available within the global environment.
 
User Action : Run the below lines
```{r}
entrop <- entr1(path)$result1 ## Run the Entropy matrix
varrow <- entr1(path)$result2 ## 
varcol <- entr1(path)$result3

```
 
 Step : 4
 
The below function is the main function to calculate the entropies between the selected variables and also to calculate the correlation between the variables.
 
User Action : Run the entr2 function

```{r}

entr2 <- function(wind){
  if((wind > varrow) | (wind > varcol)){stop("Window selected is out of range")} ## Stopping the function if the window selected is greater than the number of rows or variables present
               
entrop1 <- entrop[1:wind,1:wind]
variables <- names(entrop1) ## listing the variables which are selected using the sliding window
ent <- matrix(0,nrow = wind,ncol=1) ## Defining the empty matrix for storing the entropy values
ent1 <- matrix(0,nrow= wind,ncol=wind) ## Defining the empty matrix for storing mutual entropy values
colnames(ent1) <- variables
rownames(ent1) <- variables
gcor <- matrix(0,nrow= wind,ncol=wind) ## Defining the empty matrix for storing general correlation values
colnames(gcor) <- variables
rownames(gcor) <- variables
 
 
  for (i in 1:wind) {
 
  temp <- as.data.frame(table(entrop1[,i])) ## This is to divide into bins and calculate the frequencies of each bins. The number of bins is equal to the unique values of the variable
  temp$prob <- temp$Freq/wind ## Calculating the probabilities of each of the values
  temp$entr <- log2(temp$prob)*temp$prob*-1 ## Calculating the log2 values of the probilities
  ent[i] <- sum(temp$entr) ## Calculating the entropies of each variable                 
 
                      }

## Second For loop for calculating entropy and correlations
 
    for ( i in 1:wind-1){
               
    j = i+1
               
      for(k in j:wind){
 
        temp1 <- entrop1[,c(i,k)] ## Creating a temporary data frame with only the relevant variables
        temp2 <- as.data.frame(table(temp1))
        temp2$prob <- temp2$Freq/wind ## Calculating the probabilities of the mutual terms
        temp2$entr <- log2(temp2$prob)*temp2$prob*-1 ## Calculating the log2 values of the mutual terms
        temp3 <- sum(temp2$entr,na.rm=TRUE)
        ent1[i,k] <- ent[i] + ent[k]-temp3 ## Calculating the mutual entropy values
        ent1[k,i] <- ent1[i,k]
        gcor[i,k] <- (1-(exp(-2*ent1[i,k])))^0.5 ## Calculating the general correlation between the variables
        gcor[k,i] <- gcor[i,k]
                                               
                         }
                                                
    ent1[i,i] <- 1
    gcor[i,i] <- 1
                          }             
    ent1[wind,wind] <- 1 ## The below lines are to update the last diagonal cell with 1
    gcor[wind,wind] <- 1
 
   
  list(result1=ent1,result2=gcor)   
  
}


```

Step 5
 
The below lines have to be executed to find the number of rows and number of variables in the dataframe , so that the user can select the appropriate sliding window.
 
User Action : Run the below lines and select the sliding window 

```{r}
entr1(path)$result2 ## Run this line for selecting the number of rows
 
entr1(path)$result3 ## Run this line for finding number of columns in the dataframe

```

 
Step 6
 
The below steps executes the main function to calculate the entropy and correlation between the variables.
 
User Action : Enter the window for executing the function.
```{r}
entropy_matrix <- entr2(5)$result1 ## Run this line to get the values for the entropy matrix
 
correl_matrix <- entr2(5)$result2 ## Run this line to get the values for the correlation matrix
 
 
entropy_matrix ## displaying the entropy matrix
correl_matrix ## Displaying the correlation matrix

```

 
 





